Based on the details above, a manager is in charge of bridging the execution team and the upper management team. He has to both represent the interest of the employees and both the desires of the company. People managers have to evaluate the performance of the employees, but in order to do so he must set in advance development objectives, preferably using a \textit{SMART} template. For interacting with different employees, his attitude must be adjusted by the interlocutors \textit{PRAE} model. He must always be able to coach, but also make drastic decisions (e.g.: firing). It is important that a manager avoid conflicts with his employees and from observations it is best if the manager is a technical person from outside the team.

\subsection{Questions managers have to answer in order to interact with his reports}
\label{subsec:questions}
Managers usually have to adapt their attitude and actions in interaction with their employees. Below is a set of questions that will be used in the future work of this paper.

\begin{description}
\item[How do you adapt you attitude and language based on the interlocutors?]
\item[What is the best motivator that you use?]
\item[How do you increase the morale of your employees?]
\item[What are the tools that you rely on in your managerial activities?]
\item[How do you make the decision to fire someone?]
\item[What do you do before firing someone?]
\item[What was your biggest failure?]
\item[What lead to your greatest success?]
\item[How are you coaching your reports?]
\item[Describe the type of objectives that you set.]
\end{description}


\subsection{Future Work}
\label{subsec:future}
This report will be enhance by a study about people management in the Romanian IT industry. The study will consist of direct interviews with people managers from Bucharest based companies using the questions from \cref{subsec:questions}. Also seeing that people management and team leadership have similar responsibilities a comparison between the two of them will be conducted to see if how far apart are one from another.

People management is not yet widely used, but it is tarting to cache ground. Seeing that every person reacts differently is important for a people manager to predict how his reports will take his decisions and adjust his attitude to each and one of them. If a manager does not ask himself \textit{WHY} things are the way they are and treat only the effects of the actions, result will never be optimal.