% vim: set tw=78 sts=2 sw=2 ts=8 aw et ai:

A general guideline to creating a well working team is easy to understand but it does not represent a recipe guaranteed to succeed. Every action and decision can be challenged.

\subsection{Resistance to change}
Even if the results of a team are not good, and they know it, they will not be very open to try something new. This fact is caused by the idea that each member gives all his best to make things works, but it is still not enough. In leadership this is called \textit{unconscious incompetence}, or simpler said \textit{they are not aware of what they don't know}. Because of this the members will counter the idea of learning new skills and they will find any other excuse for the bad performances. The leader must make them aware of the alternatives that are available and their main advantages and disadvantages. Each member will try to make things work as he is accustomed from his previous experiences. The leader must ensure that the team reaches a consensus about the way that they will move forward. 
 
\subsection{Financial needs}
Financial security is considered a \textit{hygiene factor}. This means that the presence and abundance of money generates a short term motivation that usually lasts between two and three months after the salary increase, while in the same time it means that the absence of money represents an important de-motivator. In the IT the pressure the satisfy the salary expectations is enhanced by the fact that the request for software engineers is far bigger than what the universities provide. In Romania it's estimated that the requirement is of 20000 engineers while the offer is 11000 engineers per year. This means that the competitors can provide a more appealing offer. A team leader has no direct control over the salary of the team members, but his efforts to build a team are greatly impacted by a bad salary.

This two impediments are not the only ones that a team leader can face, but are the most usual ones. Other impediments that can arise are:

\begin{enumerate}
\item bad attitude from some members
\item management pressures
\item lack of resources (technical or human)
\end{enumerate}