The guidelines described above are usually applicable in the software industry, but are not applicable in all industries.
The following industries have different approaches because they face other constrains.

\subsection{Military field}
The military field is the root of the \textit{Leader-Follower} model. In the army there is always the need to have predictability over the actions of a platoon. Every soldier has to execute the same manoeuvre in sync with the rest of the platoon. Improvement and empowerment inside a formation will result in a formation loss. In case of war, the strategy is planned at a high level and the orders are cascaded down. If a middle officer tries to challenge an order in good or bad reason, the whole strategy can fail. In war times every strategy is based on precise timing and this kind of discussions will add an time overhead, thus missing a window of opportunity. In the same time the military field is based on long term contracts, soldiers not having the privilege to just quit and find another job. Also the only way to really grow is to be observed by a superior and prove potential. If this potential is seen, the soldier is proposed to follow intensive courses in simulated situations. After the courses are done, the soldier has new objectives that he has to meet in the exact way that he was taught. 

\subsection{Medical field}
In surgery the team coordination is very similar to the military field. Each surgery can be seen as a day-by-day phase. It is not the time to learn new skills by trial and error. The surgery is coordinated by a head surgeon that gives clear directions and makes all the decisions. Other doctors may help, but how and when is up to the head surgeon. The pressure is greater that in the military field, the result can be very devastating (ranging from inability to death of the patient). The learning process for surgeons is very difficult that for engineers. They can not experiment on the job. To learn a new technique, they first have to learn the theory, then they have to observe the experience doctors and then start to practice by doing the simple steps of the process while the complicated ones are done by the doctors with experience. After doing the simple steps a couple of times, the head surgeon can start to delegate more complicated steps. Compared to the software industry the delegation can not be challenging.